---
title: "HW04"
author: "Saelin"
date: "08/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading packages needed:
```{r}
library(ggplot2)
library(tidyverse)
library(gapminder)
```

```{r}
gapminder
```
## Exercise 1

** Univariate Option 1: **
1.1 Make a tibble with one row per year, and columns for life expectancy for two or more countries

```{r}
gapminder_wide <- gapminder %>%
  filter(country %in% c("China", "Taiwan", "Japan")) %>%
  pivot_wider(names_from = country,
              values_from = lifeExp,
               id_cols = year )
as_tibble(gapminder_wide)
```

1.2 Shape a scatterplot to life expectancy for one country against that of another
```{r}
gapminder_wide %>%
  ggplot(aes(China, Japan, color = year, size = 1.5)) +
  geom_point() +
  labs(title = "Life Expectancy of China vs. Japan", subtitle = "From 1952 to 2007", color = "Year") +
  guides(size = FALSE) + #remove size legend
  theme_light()
```

1.3 Re-lengthen the data
```{r}
gapminder_long <- gapminder_wide %>%
  pivot_longer(names_to ="Country",
               values_to = "Life Expectancy", 
                 cols = -year) 
as_tibble(gapminder_long)
```

## Exercise 2

** Multivariate Option 2: **

2.1 Make a tibble with one row per year, and columns for life expectancy and GDP per capita for two or more countries
```{r}

```
2.2 Re-lengthen the data
```{r}

```


## Exercise 3

3.1 For each guest in the guestlist (guest tible), add a column for email address, which can be found in email table

```{r }
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")

```

3.2 Who do we have emails for, yet are not on the guestlist?

```{r 3.2 }

```


3.3 Make a guestlist that includes everyone we have emails for (in addition to those on the original guestlist)
```{r 3.3 }

```