---
title: "HW04"
author: "Saelin"
date: "08/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading packages needed:
```{r}
library(ggplot2)
library(tidyverse)
library(gapminder)
```

The Data:
```{r}
as_tibble(gapminder)
```
## Exercise 1

** Univariate Option 1: **
1.1 Make a tibble with one row per year, and columns for life expectancy for two or more countries

```{r}
gapminder_wide <- gapminder %>%
  filter(country %in% c("China", "Taiwan", "Japan")) %>%
  pivot_wider(names_from = country,
              values_from = lifeExp,
               id_cols = year )
as_tibble(gapminder_wide)
```

1.2 Shape a scatterplot to life expectancy for one country against that of another
```{r}
gapminder_wide %>%
  ggplot(aes(China, Japan, color = year, size = 1.5)) +
  geom_point() +
  labs(title = "Life Expectancy of China vs. Japan", subtitle = "From 1952 to 2007", color = "Year") +
  guides(size = FALSE) + #remove size legend
  theme_light()
```

1.3 Re-lengthen the data
```{r}
gapminder_long <- gapminder_wide %>%
  pivot_longer(names_to ="Country",
               values_to = "Life Expectancy", 
                 cols = -year) 
as_tibble(gapminder_long)
```

## Exercise 2

** Multivariate Option 2: **

2.1 Make a tibble with one row per year, and columns for life expectancy and GDP per capita for two or more countries
```{r}
gap_multi_wide <- gapminder %>%
  filter(country %in% c("China", "Taiwan", "Japan")) %>%
  rename(GDP_PerCap = gdpPercap, Life_Expectancy = lifeExp) %>%
  pivot_wider(id_cols = year,
              names_from = country,
              values_from = c(GDP_PerCap, Life_Expectancy), 
              names_sep = " ")
as_tibble(gap_multi_wide)
```

2.2 Re-lengthen the data
```{r}
gap_multi_long <- gap_multi_wide %>%
  pivot_longer(cols = -year,
               names_to = c("Metric","country"), 
               values_to= c("Measurement"),
               names_sep = " ")
as_tibble(gap_multi_long)
  
```


## Exercise 3

The Data:
```{r }
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
as_tibble(email)
as_tibble(guest)
```

3.1 For each guest in the guestlist (guest tible), add a column for email address, which can be found in email table

```{r }
emails <- email %>%  #modify email to make it "joinable"
  rename("name" = guest) %>%   #rename guest to name 
  separate_rows("name", sep =", ")  # seperate names within one cell into separate rows

as_tibble(left_join(guest, emails))  # join the tables, adding email column to guest

```

3.2 Who do we have emails for, yet are not on the guestlist?

```{r 3.2 }
emails <- email %>%  #modify email to make it "joinable"
  rename("name" = guest) %>%   #rename guest to name 
  separate_rows("name", sep =", ")  # seperate names within one cell into separate rows

as_tibble(anti_join(emails, guest)) #filter-join the tables, keeping only rows in emails that aren't in guest

```


3.3 Make a guestlist that includes everyone we have emails for (in addition to those on the original guestlist)
```{r 3.3 }
emails <- email %>%  #modify email to make it "joinable"
  rename("name" = guest) %>%   #rename guest to name 
  separate_rows("name", sep =", ")  # seperate names within one cell into separate rows

as_tibble(full_join(guest,emails))


```



