---
title: "HW02 dplyr exploration"
output:
  html_document:
  toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load, warning = FALSE, echo = FALSE, message = FALSE}
library(tidyverse)
library(gapminder)
library(dplyr)
library(ggplot2)
library(knitr)
library(DT)
```

## Exercise 1

### Use filter() to subset the gapminder data to three countries of your choice in the 1970’s.

```{r 1.1, warning = FALSE, echo = FALSE}
subset <- filter(gapminder, year > 1969 & year < 1980) 
subset <- filter(subset, country=="Japan" | country =="Costa Rica" | country=="Egypt")
kable(subset, align="c")
```

### Use the pipe operator %>% to select “country” and “gdpPercap” from your filtered dataset in 1.1.

```{r 1.2, warning = FALSE, echo = FALSE}
subset <- filter(gapminder, year > 1969 & year < 1980) 
subset <- filter(subset, country=="Japan" | country =="Costa Rica" | country=="Egypt")
subset %>% 
  select(country, gdpPercap) %>%
kable(align="c")
```

### Filter gapminder to all entries that have experienced a drop in life expectancy.

```{r 1.3, warning = FALSE, echo = FALSE}
gapminder %>% 
  arrange(country, year) %>%
  group_by(country) %>%
  mutate(Diff_LifeExp = lifeExp - lag(lifeExp)) %>%
  filter(Diff_LifeExp < 0) %>%
datatable()
```

### Filter gapminder to contain six rows: the rows with the three largest GDP per capita, and the rows with the three smallest GDP per capita.


```{r 1.4, warning = FALSE}
gapminder %>%
  arrange(gdpPercap) %>%  # arranges in ascending order
  slice(-4:-(n()-3)) %>% # drops all rows from 4 to [(last row)-3], leaving first and last 3
  kable(align="c")
```


### Produce a scatterplot of Canada’s life expectancy vs. GDP per capita using ggplot2, without defining a new variable. 

```{r 1.5, warning = FALSE}
gapminder %>%
  filter(country == "Canada") %>%
  ggplot(aes(x=gdpPercap, y = lifeExp)) +
  geom_point() +
  scale_x_log10("GDP Per Capita") +
  ylab("Life Expectancy") + 
  ggtitle("Life Expectancy vs. GDP Per Capita in Canada") +
  theme_light()
```

##Exercise 2

### Exploring categorical (country) and quantitative (population) variables 


Population data:

```{r 2.1, warning = FALSE}
gapminder %>%
  summarize(mean_pop = mean(pop), 
            highest_pop = max(pop),
            min_pop = min(pop), 
            sd_pop = sd(pop), 
            pop_range = (range(pop))[2] - (range(pop))[1], 
            num_countries = length(unique(country))) %>% 
             kable(align="c")
```

Population by country:

```{r 2.2, warning = FALSE}
gapminder %>%
  group_by(country) %>%
  summarize(mean_pop = mean(pop), 
            highest_pop = max(pop),
            lowest_pop = min(pop), 
            sd_pop = sd(pop), 
            range = (range(pop))[2] - (range(pop))[1],
            year_highest_pop = year[pop == highest_pop],
            year_lowest_pop = year[pop == lowest_pop]) %>%
   datatable()
```

Graphing world population vs. year, where color = continent and size = GDP per capita. 
```{r 2.3, warning = FALSE}

gapminder %>%
  group_by(continent) %>%
  mutate(pop = pop/1000000000) %>%
  ggplot(aes(x=year, y=pop, color=continent, size=gdpPercap)) +
  geom_point(alpha=0.8) +
  xlab("Year") + 
  ylab(expression(paste("Population ", 10^9))) +
  theme_light()

``` 

Graphing countries with populations over 50 million in 1952

```{r 2.4, warning = FALSE}
gapminder %>%
  filter(pop > 50000000 & year == 1952) %>%
   mutate(pop = pop/1000000000) %>%
  ggplot(aes(x=country, y=pop, fill=country)) +
  geom_bar(stat="identity") +
  xlab("Country") + 
  ylab(expression(paste("Population (", 10^9, ") in 1952"))) +
  theme_light() +
  theme(axis.text.x=element_text(angle=45, vjust=0.7))

``` 

## Exercise 3

### Make two plots that have some value to them

1. Scatter plot: Quantitative variables year, population, life expectancy.

```{r 3.1, warning = FALSE}
gapminder %>%
  mutate(pop=pop/1000000000) %>%
  ggplot(aes(x=year, y=lifeExp, size = pop, color=continent)) +
  geom_point() +
  theme_minimal() + 
  ylab("Life Expectancy") +
  xlab("Year")+
  scale_size_continuous(name=expression(paste("Population (X ", 10^9,")"))) +
  scale_color_discrete(name="Continent")
```


2. Non-scatter plot: Life expectancy vs year for countries in Africa 
```{r 3.2, warning = FALSE}
  gapminder %>%
    filter(continent == "Africa") %>%
    ggplot(aes(x=year, y=lifeExp, color=country)) +
    geom_line() +
    ylab("Life Expectancy") + 
    xlab("Year") +
    scale_x_continuous(breaks = seq(1952, 2007, by = 5)) +
    theme_minimal() +
    theme(axis.text.x=element_text(angle=45, vjust=0.5, size=7))

```

## Optional: Recycling

### Presumably the analyst’s intent was to get the data for Rwanda and Afghanistan. Did they succeed? Why or why not? If not, what is the correct way to do this?

```{r, warning = FALSE}
filter(gapminder, country == c("Rwanda", "Afghanistan"))
```

This way does not suceed in selecting all rows with these countries, this prints only half of the number of rows corresponding to these countries.

The correct to get all the data from these countries would be:
```{r, warning = FALSE}
filter(gapminder, country %in% c("Rwanda", "Afghanistan"))
```

an alternative way:
```{r, warning = FALSE}
filter(gapminder, country == "Rwanda" | country =="Afghanistan")
```

