---
title: "HW03"
author: "Saelin Bjornson"
output: 
  html_document
---

<style>
  .col2 {
    columns: 2 200px;         /* number of columns and width in pixels*/
  }
  .col3 {
    columns: 3 100px;
  }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load, include=FALSE}
library(dplyr)
library(gapminder)
library(ggplot2)
library(knitr)
library(DT)
```
## The data
# I will be working with the gapminder data set:


```{r Data}
datatable(gapminder)

```



# Task Option 1
## What countries had a life expectancy below the global median for each year?

```{r Task 1.1}
gapminder %>%
  group_by(year) %>%
  mutate(median_lifeExp = median(lifeExp)) %>%
  filter(lifeExp < median_lifeExp) %>%
  select(country, year, lifeExp, median_lifeExp) %>%
  arrange(year, lifeExp, country) %>%
  rename(Country = country, Year = year, "Life Expectancy" = lifeExp, "Global Median Life Expectancy" = median_lifeExp) %>%
  as_tibble() %>%
  datatable()
```

## To summarize, how many countries on each contient had a life expectancy below the global median for each year?

<div class="col2">
```{r Task 1.2}
gapminder %>%
  group_by(year) %>%
  mutate(median_lifeExp = median(lifeExp)) %>%
  group_by(continent) %>%
  filter(lifeExp < median_lifeExp) %>%
  mutate(num_countries = length(unique(country))) %>%
  select(continent, year, median_lifeExp, num_countries) %>%
  rename(Year = year, "Global Median Life Expectancy" = median_lifeExp, "Number of Countries" = num_countries) %>%
  as_tibble() %>%
  datatable()
```

```{r Task 1.3}
gapminder %>%
  group_by(year) %>%
  mutate(median_lifeExp = median(lifeExp)) %>%
  group_by(continent) %>%
  filter(lifeExp < median_lifeExp) %>%
  mutate(num_countries = length(unique(country))) %>%
  select(continent, year, median_lifeExp, num_countries) %>%
  ggplot(aes(year, num_countries, fill = continent)) +
  geom_bar(position="stack", stat="identity")
```


</div>



# Task Option 2
## What is the maximum and minimum GDP per capita for all continents?
```{r Task 2}

```


# Task Option 4
## Reporting the mean life expectancy, weighted by life expenctancyand trimmed 10%

```{r Task 4}

```
<div class="clearer"></div>



